---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mitools

<!-- badges: start -->
<!-- badges: end -->

mitools is an R package designed to help users create sets of variables based on a mutual information approach. In this context, a set is a collection of distinct elements (e.g., variables) that can also be treated as a single entity. Mutual information, a concept from probability theory, quantifies the dependence between two variables by expressing how much information about one variable can be gained from observing the other.

## Installation

You can install mitools like with the following code snippet:

``` r
devtools::install_github('nicolasleenaerts/mitools')
```

## Pairing variables

You can create sets of variables with the *pairmi* function. It takes a dataframe of variables and pairs them until a certain depth (i.e., a maximum number of elements). For each set of variables, the mutual information between the variables is calculated, after which a G-statistic is calculated and evalauted for its significance under a chi-squared distribution with a prespecified alpha. Alternatively, the user can set a certain threshold of mutual information to determine the significance of the sets. 

```{r example 1, results='hide',message=FALSE}
# Loading the package, which automatically also downloads the example data (misimdata)
library(mitools) 

# Pairing variables
results = pairmi(misimdata[,2:6])

# View data with sets added
View(results$expanded.data)
```

```{r table 1,echo=FALSE,results='asis'}
knitr::kable(results$expanded.data[c(1:5),],caption = 'Table 1. Expanded Data',align = c('c'))
```

```{r example 2,results='hide',message=FALSE}
# View information on the sets
View(results$pairs)
```

```{r table 2,echo=FALSE,results='asis'}
knitr::kable(results$pairs,caption = 'Table 2. Information on sets',align = c('c'))
```


