---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# setweaver

*setweaver* is an R package designed to help users create sets of variables based on a mutual information approach. In this context, a set is a collection of distinct elements (e.g., variables) that can also be treated as a single entity. Mutual information, a concept from probability theory, quantifies the dependence between two variables by expressing how much information about one variable can be gained from observing the other.

## Authors
[Aaron Fisher](https://psychology.berkeley.edu/people/aaron-fisher)\
[Nicolas Leenaerts](https://nicolasleenaerts.github.io/)

## Installation

You can install *setweaver* with the following code snippet:

``` r
devtools::install_github('nicolasleenaerts/setweaver')
```

## Pairing variables

You can create sets of variables using the *pairmi* function, which takes a dataframe of variables and pairs them up to a specified maximum number of elements. For each set, the mutual information between the variables is computed, followed by the calculation of a G-statistic. This statistic is then evaluated for significance based on a chi-squared distribution with a predefined alpha level. Alternatively, users can specify a mutual information threshold to determine the significance of the sets.

```{r example 1, results='hide',message=FALSE}
# Loading the package, which automatically also downloads the example data (misimdata)
library(setweaver) 

# Pairing variables
results = pairmi(misimdata[,2:11],alpha = 0.05,n_elements = 5)

# View data with sets added
View(results$expanded.data)
```

```{r table 1,echo=FALSE,results='asis'}
knitr::kable(results$expanded.data[c(1:5),],caption = 'Table 1. Expanded Data',align = c('c'))
```

```{r example 2,results='hide',message=FALSE}
# View information on the sets
View(results$sets)
```

```{r table 2,echo=FALSE,results='asis'}
knitr::kable(results$sets,caption = 'Table 2. Information on sets',align = c('c'))
```

## Evaluating sets

Once the sets are created with the *pairmi* function , you can assess their relationship with a specific outcome using the *probstat* function. This function employs k-fold cross-validation to compute parameters such as conditional probability, conditional entropy, and the odds ratio of the outcome given a particular set. Additionally, Fisher's exact test is performed to determine whether the outcome is significantly more likely to occur in the presence of a given set of variables.

```{r example 3, results='hide',message=FALSE}
# Evaluating the sets
evaluated_sets = probstat(misimdata$y,results$expanded.data[,results$sets$set],nfolds = 5)
```

```{r table 3,echo=FALSE,results='asis'}
knitr::kable(evaluated_sets[c(1:5),],caption = 'Table 3. Evaluated sets',align = c('c'))
```

## Visualizing sets

You can visualize the sets created with the *pairmi* function using the *setmapmi* function. This function generates a setmap, which illustrates the composition of sets by showing which original variables are included in sets of a given size.


```{r example4,dev='png', fig.show='hide',message=FALSE}
# Visualizing the sets
setmapmi(results$original.variables,results$sets,n_elements = 2)
```

![Plot 1. Setmap of sets that consist of 2 elements](`r knitr::fig_chunk('example4', 'png')`)
